\chapter{Publications and other research results}
\label{ch:appendix-publications}


\section*{Publications}

{\noindent \bfseries Journals}
\begin{itemize}
    \item Nieves R. Brisaboa, Antonio Fari\~na, Daniil Galaktionov, and Tirso V. Rodeiro.
	Trippy: a GIS to visualize aggregated users' trips data built on a compact representation.
	Draft to submit.

    \item Nieves R. Brisaboa, Antonio Fari\~na, Daniil Galaktionov, Tirso V. Rodeiro, and M. Andrea Rodr\'iguez.
	Improved structures to solve aggregated queries for trips over public transportation networks.
	Draft to submit.

	\item Nieves R. Brisaboa, Antonio Fari\~na, Daniil Galaktionov, and M. Andrea Rodr\'iguez.
	A Compact Representation for Trips over Networks built on self-indexes.
	In \textit{Information Systems}, vol 78 (2018), pages 1--22.
\end{itemize}


{\noindent \bfseries International conferences}

\begin{itemize}
	\item Nieves R. Brisaboa, Antonio Fari\~na, Daniil Galaktionov, Tirso V. Rodeiro, and M. Andrea Rodr\'iguez.
	New structures to solve aggregated queries for trips over public transportation networks.
	In \textit{Proceedings of the 25th International Symposium on String Processing and Information Retrieval (SPIRE)} (2018), pages 88--101.
	
	\item Nieves R. Brisaboa, Antonio Fari\~na, Daniil Galaktionov, and M. Andrea Rodr\'iguez.
	Compact Trip Representation over Networks.
	In \textit{Proceedings of the 23rd International Symposium on String Processing and Information Retrieval (SPIRE)} (2016), pages 240--253.
\end{itemize}

\section*{International research stays}

\begin{itemize}

    \item {\itshape May, 2018 - August, 2018.} Research stay at Universidad de Concepci\'on,
Departamento de Computaci\'on (Concepci\'on, Chile).

	\item {\itshape September, 2018 - October, 2018.} Research stay at Universidad de Chile,
Departamento de Ciencias de la Computaci\'on (Santiago, Chile).

\end{itemize}
